On Sample Size Needed for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2024-01-09 Manuscript/manuscript.tex
final smoothing

The coverage rates of the CIs were used to compare the performance of CIs. Let
$\hat\theta_{L, r}$ and $\hat\theta_{U, r}$ be the lower and upper bound,
respectively, for the confidence interval constructed for each replicate
$r \in \{1, \ldots, R\}$, where $R$ is the number of replicates. Then
$cov = \#\{\hat\theta_{L, r} < \theta < \hat\theta_{U, r} \}/R$, for
$r \in \{1, \ldots, R\}$. If a CI method is valid, then the coverage rate is
expected to match the nominal level. Because it is unlikely for the coverage to 
exactly match the nominal level, we can construct a 95\% Clopper-Pearson
exact CI of the 
coverage rate,\citep{clopper1934use}
which is an estimate of a proportion with $R = 10,000$. We used the
R \textsl{PropCIs} package to achieve this.\citep{PropCIs}
The choice of
Clopper-Pearson was motivated by the Wald 
interval's poor coverage as the proportion approaches 0 or 1, as noted by 
\citet{brown2001interval}, although when we tried Wald intervals, the coverage
rate intervals
did not appear to have large differences. 
If the 
proportion 0.95 
is included in the interval, the block bootstrap method is likely performing 
well. If all values in the interval are below 0.95, the results would suggest 
that the method either is providing inaccurate estimation, is underestimating 
the process' variability, or a combination of both. If all values in the 
interval are above .95, the results suggest that the method is overestimating 
the process' variability. 
\textbf{Figure~\ref{fig:norm1}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a marginal standard normal distribution using block bootstrap
with $l = \lceil n^{1/3} \rceil$, generated using the 
R \textsl{ggplot2} package.\citep{ggplot2}

@@

For a marginal exponential  distribution, the empirical coverage rates for 
$\mu$, $\sigma_w$, and 
the lag-1 autocorrelation coefficient $\rho$ using block bootstrap
with $l = \lceil n^{1/3} \rceil$, as well as 95\%
confidence intervals of the real coverage are displayed in 
\textbf{Figure~\ref{fig:exp1}}.

@@

\jy{Are we reporting the C-P intervals? Clarify the desciption and
  cite the reference.}
\mc{addressed}

%situations.
%Figure~\ref{fig:norm_mu_1_cp} shows the same coverage rates using exact 
%Clopper-Pearson intervals to verify that it is the choice of Wald interval
%does not cause undercoverage of the coverage rate. Indeed, the widths of the
%exact proportion CIs do not seem to differ heavily from the Wald-type CIs.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/norm_phi_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for 
    samples of $n \in \{800, 1600\}$ ($l = \lceil n^{1/3} \rceil$)
    for an AR(1) process with $\phi = 0.4$. For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:npi}
\end{figure}

@@

The
plot and results for $l = \lceil 2n^{1/3} \rceil$ are included in the Appendix.
Performance generally seems to be better for $l = \lceil n^{1/3} \rceil$, but
importantly, the changes in performance when varying other parameters do not
change when a different block size is used.

@@

Additionally, a set of 50 CIs are displayed for each of the percentile, 
BC, and recentered percentile approaches for exponentially distributed
samples of $n \in \{800, 1600\}$ with lag-1 autocorrelation coefficient 0.355
($\phi = 0.4$) in Figure~\ref{fig:eri}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/exp_phi_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for the
    lag-1 autocorrelation coefficient 0.355 ($\phi = 0.4$)
    of a stationary series with marginal exponential distribution
    obtained by transforming an AR(1) process with $\phi = 0.4$ with
    sample size $n \in \{800, 1600\}$ ($l = \lceil n^{1/3} \rceil$). For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:eri}
\end{figure}

It appears that a greater sample size is generally required for the bootstrap
CIs to cover the mean and standard deviation parameters in the exponential
margin case than in the normal margin case. In contrast to the marginal normal
distribution, where standard normal, Student's $t$, and 
recentered percentile CI $\phi$ coverage rates were greater for stronger 
dependences 
regardless of the sign, $\rho$ coverage rates for these CIs seem to be greater 
only for 
stronger negative dependencies and lesser for stronger positive dependences. 
\mc{What could be an explanation for this?}
\jy{I don't follow this. Are you talking about mu only? Even for mu,
  the statement does not look right.} \mc{for lag-1 autocorrelation}
\jy{This would invalidate our explanation using the effective sample
  size. Let's tune it down or remove it.}
\mc{this observation was for lag-1 autocorrelation. does it still invalidate 
our effective sample size explanation?}
However, the other 
trends 
and patterns
discussed in the main text regarding the performance of various methods and
diverse parameters remain consistent. For example, Student's $t$ confidence
intervals still exhibit higher coverage rates in comparison to alternative
methods. Performance continues to be more favorable when temporal dependence is
negative rather than positive. Of particular significance, the percentile, BC,
and BCA confidence intervals still display a decline in coverage accuracy for
the lag-1 autocorrelation coefficient as sample size increases as demonstrated
in Figure~\ref{fig:eri}. Both the percentile and BC intervals persist in
manifesting the same bias issue. On the other hand, the recentered percentile
confidence interval continues to be effective in estimating the temporal 
dependence due to the
inherent unbiasedness of the original point estimator. In sum, 
for the most part, the findings for
series that are marginally exponentially distributed closely mirror those
attained for series that are marginally normally distributed.

@@

Block bootstrap is a useful method for estimating parameters of a time series,
from simple parameters like the mean to more complicated temporal dependence
factors. We know theoretically that the block bootstrap procedure will cover a 
parameter of a time series at the nominal level given an infinitely large 
sample,\citep{calhoun2018} so the goal for this study was to find the smallest 
finite sample length 
$n$ of a time series in order for the block bootstrap procedure to recover its 
associated parameters at an acceptable rate. Our analysis relies on the 
assumption that there is a size $n$ large enough for the method to work: that 
is, the method's performance improves as $n$ increases. Out of the six types of 
intervals used in this study, this assumption was found to hold true with 
respect to estimating $\phi$ only for standard normal, Student's $t$, and recentered 
percentile CIs, whereas percentile, BC, and BCA intervals exhibited coverage 
deterioration as $n$ increased. The percentile CI's coverage deterioration can 
be attributed to bias that is not corrected as $n$ increases. Specifically, as 
$n$ increases, the width of the CI decreases, but because the percentile CI 
underestimates $\phi$, the coverage decreases. The BC CI seems to correct the 
bias, but the width of the CI seems to be too short. The acceleration factor 
of the BCA CI seems to fail, as the width of the CI seems to be too
short.
\jy{Inconsistent means not consistent and consistency is a desired
  property of an estimator. When we talk about width of a CI, what
  does consistency mean? Do an throughout check, please.}
\mc{I was basically trying to say that the widths of the CI vary more 
for BC CIs than for percentile or recentered CIs, I think it may be simpler to
just say that they are shorter}

One of the goals of this study was to provide some practical recommendations for
necessary sample sizes when using block bootstrap to estimate the parameters of
serially dependent data. When using Student's $t$ intervals and the
marginal
\jy{Recall that t is not different from normal when sample size is
  large. So the recommendation is to do t always since it performs
  better for small samples.} \mc{addressed}
distribution and temporal dependence is 
unknown, the results of this study suggest that $n \geq 1600$ may be necessary 
for common practice to estimate $\mu$, whereas $n > 3200$ may be necessary
to estimate the standard deviation. Student's $t$ is always preferable to
Standard Normal CIs
as they performs better for smaller sample sizes and performs as good or better
for larger sample sizes.
\jy{Don't discuss these. The recommended method is regardless of what
  the dependence struture and marginald istribution are. That's why we
  did the study. Just gave recommendations based on both
  scenarios. Rewrite this paragraph.}
\mc{addressed}
\jy{These statements are not to the point. Let's not discuss the case
  where phi is know or not. They are not known. We want the method to
  be robust to phi and the true marginal distribution.}
\mc{addressed}
Lastly, to estimate 
$\phi$, $n \geq 100$ using the Student's $t$ method may be sufficient under
a marginal standard normal distribution, whereas $n \geq 1600$ may be required
under under a marginal exponential distribution. Further
investigation may be necessary to see if there are other percentile-based 
interval corrections
that fix the coverage deterioration problem for~$\phi$.

@@

\textbf{Figure~\ref{fig:exp2}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a marginal exponential distribution using block bootstrap
with $l = \lceil 2n^{1/3} \rceil$.

@@

For a stationary series with a marginal unit exponential distribution, altering 
the 
block size results in the same 
changes in performance of different CIs.

>>>>

The coverage rates of the CIs were used to compare the performance of CIs. Let
$\hat\theta_{L, r}$ and $\hat\theta_{U, r}$ be the lower and upper bound,
respectively, for the confidence interval constructed for each replicate
$r \in \{1, \ldots, R\}$, where $R$ is the number of replicates. Then
the empirical coverage rate is
$ \sum_{r=1}^{R} I\{\hat\theta_{L, r} < \theta < \hat\theta_{U, r}
\}/R$, where $I(\cdot)$ in the indicator function.
If a CI method is valid, then the coverage rate is
expected to match the nominal level. Because it is unlikely for the coverage to 
exactly match the nominal level, we can construct a 95\% Clopper-Pearson
exact CI of the coverage rate,\citep{clopper1934use}
which is an estimate of a proportion with $R = 10,000$. We used the
R \textsl{PropCIs} package to achieve this.\citep{PropCIs}
The choice of Clopper-Pearson was motivated by the Wald 
interval's poor coverage as the proportion approaches 0
or~1,\citep{brown2001interval}
although when we tried Wald intervals, the coverage
rate intervals did not appear to have large differences. If the proportion 0.95
is included in the interval, the block bootstrap method is likely performing 
well. If all values in the interval are below 0.95, the results would suggest 
that the method either is providing inaccurate estimation, is underestimating 
the process' variability, or a combination of both. If all values in the 
interval are above .95, the results suggest that the method is overestimating 
the process' variability. \textbf{Figure~\ref{fig:norm1}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a marginal standard normal distribution using block bootstrap
with $l = \lceil n^{1/3} \rceil$, generated using the 
R \textsl{ggplot2} package.\citep{ggplot2}

@@

@@

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/norm_phi_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for 
    samples of size $n \in \{800, 1600\}$ ($l = \lceil n^{1/3} \rceil$)
    for the lag-1 autocorrelation of an AR(1) process with $\phi = 0.4$. For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:npi}
\end{figure}

@@

The results for $l = \lceil 2n^{1/3} \rceil$ are reported in the Appendix.
Performance generally seems to be better for $l = \lceil n^{1/3} \rceil$, but
importantly, the patterns in performance when varying other parameters
appear to be robust to the different block size.

@@

For the scenario of marginal exponential distribution, the empirical
coverage rates for $\mu$, $\sigma_w$, and 
the lag-1 autocorrelation coefficient $\rho$ using block bootstrap
with $l = \lceil n^{1/3} \rceil$, as well as 95\%
confidence intervals of the real coverage are displayed in 
\textbf{Figure~\ref{fig:exp1}}. 
Additionally, a set of 50 CIs are displayed for each of the percentile, 
BC, and recentered percentile approaches for exponentially distributed
samples of $n \in \{800, 1600\}$ with lag-1 autocorrelation coefficient 0.355
($\phi = 0.4$) in \textbf{Figure~\ref{fig:eri}}.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/exp_phi_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for the
    lag-1 autocorrelation coefficient $\rho = 0.355$ ($\phi = 0.4$)
    of a stationary series with marginal exponential distribution
    obtained by transforming an AR(1) process with $\phi = 0.4$ with
    sample size $n \in \{800, 1600\}$ ($l = \lceil n^{1/3} \rceil$). For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:eri}
\end{figure}

It appears that a greater sample size is generally required for the bootstrap
CIs to cover the mean and standard deviation parameters in the exponential
margin case than in the normal margin case. However, the other trends 
and patterns discussed regarding the performance of various methods and
diverse parameters remain unchanged. For example, Student's $t$ confidence
intervals still exhibit higher coverage rates in comparison to alternative
methods. Performance continues to be more favorable when temporal dependence is
negative rather than positive. Of particular significance, the percentile, BC,
and BCA confidence intervals still display a decline in coverage accuracy for
the lag-1 autocorrelation coefficient as sample size increases as demonstrated
in Figure~\ref{fig:eri}. Both the percentile and BC intervals persist in
manifesting the same bias issue. On the other hand, the recentered percentile
confidence interval continues to be effective in estimating the temporal 
dependence due to the
inherent unbiasedness of the original point estimator. In sum, 
for the most part, the findings for
series that are marginally exponentially distributed closely mirror those
attained for series that are marginally normally distributed.

@@

Block bootstrap is a useful method for estimating parameters of a time series,
from simple parameters like the mean to more complicated temporal dependence
factors. We know theoretically that the block bootstrap procedure will cover a 
parameter of a time series at the nominal level given an infinitely large 
sample,\citep{calhoun2018} so the goal for this study was to find the smallest 
finite sample length~$n$
of a time series in order for the block bootstrap procedure to recover its 
associated parameters at an acceptable rate. Our analysis relies on the 
assumption that there is a size $n$ large enough for the method to work: that 
is, the method's performance improves as $n$ increases. Out of the six types of 
intervals used in this study, this assumption was found to hold true with 
respect to estimating $\phi$ only for standard normal, Student's $t$, and recentered 
percentile CIs, whereas percentile, BC, and BCA intervals exhibited coverage 
deterioration as $n$ increased. The percentile CI's coverage deterioration can 
be attributed to bias that is not corrected as $n$ increases. Specifically, as 
$n$ increases, the width of the CI decreases, but because the percentile CI 
underestimates $\phi$, the coverage decreases. The BC CI seems to correct the 
bias, but the width of the CI seems to be too short. The acceleration factor 
of the BCA CI seems to fail, as the width of the CI seems to be too
short.

One of the goals of this study was to provide some practical recommendations for
necessary sample sizes when using block bootstrap to estimate the parameters of
serially dependent data. When using Student's $t$ intervals and the
marginal distribution and temporal dependence is 
unknown, the results of this study suggest that $n \geq 1600$ may be necessary 
for common practice to estimate $\mu$, whereas $n > 3200$ may be necessary
to estimate the standard deviation. Student's $t$ is always preferable to
Standard Normal CIs
as they performs better for smaller sample sizes and performs as good or better
for larger sample sizes. Lastly, to estimate lag-1 autocorrelation,
$n \geq 100$ using the Student's $t$ method may be sufficient under
a marginal standard normal distribution, whereas $n \geq 1600$ may be required
under under a marginal exponential distribution. Further
investigation may be necessary to see if there are other percentile-based 
interval corrections that fix the coverage deterioration problem for~$\phi$.

@@

@@

\textbf{Figure~\ref{fig:exp2}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a marginal exponential distribution using block bootstrap
with $l = \lceil 2n^{1/3} \rceil$. Again, altering the block size results in the same
changes in performance of different CIs as those in the scenario of
marginal normal distribution.

### 2024-01-09 Manuscript/reply.tex
final smoothing

\usepackage{color}
\definecolor{darkblue}{rgb}{0.1, 0.2, 0.6}
\usepackage{xcolor}
\newcommand{\jy}[1]{\textcolor{red}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{blue}{(EDS: #1)}}
\newcommand{\mc}[1]{\textcolor{green}{(MC: #1)}}

@@

\mc{for exponential, the first-order autocorrelation coefficient
is not the same as $\phi$, should I define $\rho$ as the first-order 
autocorrelation coefficient in the manuscript so that I can use the greek letter
in the plot}
\jy{Yes.}
\mc{addressed}

@@

Thank you for pointing this out. We tried both Wald intervals and exact 
Clopper-Pearson intervals. Although we did not find a major difference between
the two, we have switched to Clopper-Pearson intervals to avoid the potential
undercoverage. We have discussed this in the third paragraph of the Marginal 
Standard Normal 
Distribution subsection of the Design section, and we have provided a reference
for Clopper-Pearson intervals \citep{clopper1934use} as well as the package we 
used \citep{PropCIs}.

\eds{I don't think this answers the question.  The reviewer is saying that 0.95 
is close to the boundary of 1, so the undercoverage observed intervals
(especially when phi -0.4) may be due, at least in part, to the flaws 
of the asymptotic Wald interval you are using. 
\@ Jun: do you think he should try exact Clopper-Pearson
intervals for at least a few cases of undercoverage to verify it is not the Wald
interval that is causing the undercoverage, then comment on this in at least the
reply (and possibly also in the main paper)?}
\mc{plot\_norm\_mu\_1\_cup.pdf (exact Clopper-Pearson) seems to have about the same 
widths for high 
proportions
(especially for phi = -0.4) as {plot\_norm\_mu\_1.pdf} (Wald-Type)}

\jy{Are there big differences? We could report the CP CIs in the paper.}
\mc{I included it in the paper but they do not seem too different}
\jy{Point out the location of the changes in the text.}
\mc{addressed}

@@

Thank you for the suggestion. However, because of the deterioration for $\phi$, 
we found it hard to fix the limits on the 
y-axis without hurting the visualization for the other rows. The other reviewer
suggested that we put the 6 figures in a column in 
a single plot.
Figure~\ref{fig:alt2_norm1} is provided as an example of this with fixed limits:

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/alt2_plot_norm1}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$, the marginal standard deviation $\sigma_x$, and
    the first-order autocorrelation coefficient $\phi$ of an AR(1) process with 
    a marginal standard 
    normal distribution with AR coefficient
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:alt2_norm1}
\end{figure}

%\jy{I'm interested in seeing the figures overlaid as in
%  Figure~\ref{fig:mu}.}
\mc{addressed}\jy{Put them overlaid; three plots in one with facet
  grid as the last one.}
  \mc{addressed. I think the last one looks better}

@@

Thank you for this comment.
A possible explanation for this is that if a 
stationary series has a positive 
autocorrelation, the effective sample
size is decreased, whereas is a series has a negative autocorrelation, the
effective sample size is increased. Additionally, this seems to have a 
greater effect on the the estimation of the location parameter versus that
of the scale parameter or temporal dependence parameter. We have included a 
comment at the end of the paragraph on estimation of~$\mu$, and we provided a 
reference for effective sample size \citep{geyer2011introduction}.

@@

Thank you for pointing this out. The original sentence has been reworded as you
suggested to make it more clear.

@@

\jy{Add a discussion in the results section, referring to detailed
  results in the appendix. Point out the location of this discussion
  in this reply.}
\mc{addressed}

@@

Thank you for the comment. We only
used serial dependences as strong as 0.4, because we only seek to 
establish the 
general trend as the strength of the autocorrelation 
increases, and how it varies depending on the sign of the autocorrelation and 
the parameter of interest. This justification has been added to the first
paragraph of the Normal section of the Design 
section. We have also added a deliberation (third paragraph in the Discussion) 
regarding what would
be expected for higher dependence levels.
\jy{give the location of the paragraph in the text.}
\mc{addressed}

>>>>

\usepackage[dvipsnames]{color}
\definecolor{darkblue}{rgb}{0.1, 0.2, 0.6}

\newcommand{\jy}[1]{\textcolor{red}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{blue}{(EDS: #1)}}
\newcommand{\mc}[1]{\textcolor{green}{(MC: #1)}}

@@

@@

Thank you for pointing this out. We tried both Wald intervals and exact 
Clopper-Pearson intervals \citep{clopper1934use}, the latter of which
as implemented in the R package \texttt{PropCIs}.\citep{PropCIs}
No major difference between the two was found, which is expected as we
have a large number replication (10,000). Nonetheless, we have
switched to reporting Clopper-Pearson intervals to avoid the potential
undercoverage and discussed this in the third paragraph of the Marginal 
Standard Normal 
Distribution subsection of the Design section.
\jy{Quanlity control the bib entries.}

@@

As suggested by the other reviewer, we now overlay the comparison of 6
CIs within the same panels. Each row reports the results for one
parameter, with a shared scale on the y-axis. The number of figures
have been greatly reduced and the comparison made easier.

@@

Thank you for this comment.
A possible explanation for this is that if a stationary series has a positive
autocorrelation, the effective sample size is decreased, whereas is a
series has a negative autocorrelation, theeffective sample size is
increased.\citep{geyer2011introduction}
Additionally, this seems to have a 
greater effect on the the estimation of the location parameter versus that
of the scale parameter or temporal dependence parameter. We have included a 
comment at the end of the paragraph on estimation of~$\mu$.

@@

Thank you for pointing this out. The sentence has been rewphrased as
suggested.

@@

@@

Thank you for the comment. We only
used serial dependences as strong as 0.4, because we only seek to 
establish the 
general trend as the strength of the autocorrelation 
increases, and how it varies depending on the sign of the autocorrelation and 
the parameter of interest. This justification has been added to the first
paragraph of the Normal section of the Design 
section. We have also added a deliberation (third paragraph in the Discussion) 
regarding what would be expected for higher dependence levels.
